Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19

### 2023-04-21 manuscript/manuscript.qmd
Add map and list of countries

```{r libraries-data}
library(tidyverse)
library(scales)
library(lubridate)
library(tidybayes)
library(marginaleffects)
library(patchwork)
library(colorspace)
library(modelsummary)
library(kableExtra)
library(gt)
library(targets)

@@

tar_load(c(policies_plot_data, human_rights_plot_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))

@@

@@

Between January 2020 and June 2021, 23 states issued a total of 114 ICCPR derogations. [ INSERT MAP ABOUT HERE ] Of derogations that detailed specific obligation suspension lengths, the average length was 41 days. Descriptively, we see states across multiple regions derogated with notable concentration in South America, which has a robust history of engaging with human rights law. We also note that it is not necessarily the 'usual suspects' of strong democracies derogating.  Notably, Nordic states, which typically are very active in human rights law engagement did not issue any derogations during this period.

@@

In all model specifications, we include new and cumulative COVID-19 cases and deaths as controls. In our second hypothesis, we add controls for previous ICCPR derogations or other actions, along with V-Dem indices for the respect of rule of law, civil liberties, and civil society.

>>>>

```{r libraries-data}
library(tidyverse)
library(scales)
library(sf)
library(lubridate)
library(tidybayes)
library(marginaleffects)
library(patchwork)
library(colorspace)
library(modelsummary)
library(kableExtra)
library(gt)
library(targets)

@@

tar_load(c(policies_plot_data, human_rights_plot_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))
tar_load(c(derogation_map_data, derogation_count))

@@

```{r make-var-lists-and-tbls}
derog_detail <- derogation_count %>% 
  filter(derogations > 0) %>% 
  summarize(countries = n(),
            total = sum(derogations)) %>% 
  as.list()

map_data_nice <- derogation_count %>% 
  filter(derogations > 0) %>% 
  arrange(desc(derogations)) %>% 
  mutate(Country = countrycode::countrycode(iso3, "iso3c", "country.name")) %>% 
  select(Country, Derogations = derogations)
```

@@

```{r fig-derogation-map, fig.width=6.5, fig.height=3.5}
#| include: true
#| echo: false
#| fig-cap: "Countries that derogated from the ICCPR in order to undertake pandemic-related emergency measures"
ggplot() +
  geom_sf(data = derogation_map_data, aes(fill = derogations_1plus),
          linewidth = 0.1, color = "white") +
  coord_sf(crs = st_crs("+proj=robin"), datum = NA) +
  scale_fill_binned(low = colorspace::lighten(clrs[1], amount = 0.8), 
                    high = clrs[1], na.value = "grey85",
                    breaks = c(1, 5, 10, 15, 20),
                    limits = c(1, 20),
                    name = "Derogations filed: ",
                    guide = guide_colorsteps(barwidth = 7, barheight = 0.4,
                                             title.vjust = 1)) +
  theme_pandem() +
  theme(panel.border = element_blank(),
        legend.title = element_text(face = "bold"),
        legend.text = element_text(size = rel(0.6)),
        legend.justification = "center")
```

```{r tbl-derogation-countries}
#| tbl-cap: Countries that derogated from the ICCPR in order to undertake pandemic-related emergency measures 
#| include: true
#| echo: false
bind_cols(dplyr::slice(map_data_nice, 1:12),
          add_row(dplyr::slice(map_data_nice, 13:n()))) %>% 
  kbl(booktabs = TRUE, linesep = "\\addlinespace[0.5em]",
      col.names = rep(c("Country", "Derogations"), 2)) %>% 
  kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA))
```

Between January 2020 and June 2021, `r derog_detail$countries` states issued a total of `r derog_detail$total` ICCPR derogations (see @fig-derogation-map and @tbl-derogation-countries). Of derogations that detailed specific obligation suspension lengths, the average length was 41 days. Descriptively, we see states across multiple regions derogated with notable concentration in South America, which has a robust history of engaging with human rights law. We also note that it is not necessarily the "usual suspects" of strong democracies that derogate. Notably, Nordic states—which typically are very active in human rights law engagement—did not issue any derogations during this period.

@@

In all model specifications, we include new and cumulative COVID-19 cases along with indicators for ICCPR derogations or other actions, and V-Dem indices for the respect of rule of law, civil liberties, and civil society.
